Analysis of observational data, to calculate the Delayed-Tau model parameters

### 2025-02-04 thesis.qmd
delayed-tau vs lilly-madau

$$
\text{SFRD} = \frac{1}{V}\sum^N_{i=1}\text{SFR}_{o,i},
$$ where $V$ is the comoving volume [^2] of a region of the universe, $N$ is the number of galaxies of the regions and SFR$_{0,i}$ is the current SFR of each galaxy $i$.

@@

$$
1+z = \sqrt{\frac{1+v/c}{1-v/c}}
$$ {#eq-z-rel} For small $v/c$ we can rewrite @eq-z-rel, as:

@@

Using the delayed τ model, we compute $\tau$, $t_{\text{sf}}$, and normalization constant $A_{\text{del}}$ with:

@@

where $\text{SFR}_0$ is given in the catalogs.

@@

To begin with the morphological types of the galaxies we first need to understand possible incompatabilities of the given types of each catalog.

::: {#fig-TTypes layout-ncol=2}
![Before the clipping of the transient types](compare/quickplots/Hubble%20Morphology.png)

@@

Scatter plots and the residuals of the Hubble Numerical morphological types
:::

>>>>

$$
\text{SFRD} = \frac{1}{V}\sum^N_{i=1}\text{SFR}_{o,i},
$$

where $V$ is the comoving volume [^2] of a region of the universe, $N$ is the number of galaxies of the regions and SFR$_{0,i}$ is the current SFR of each galaxy $i$.

@@

$$
1+z = \sqrt{\frac{1+v/c}{1-v/c}}
$$ {#eq-z-rel}

For small $v/c$ we can rewrite @eq-z-rel, as:

@@

Using the delayed-$\tau$ model, we compute $\tau$, $t_{\text{sf}}$, and normalization constant $A_{\text{del}}$ with:

@@

where $\text{SFR}_0$ is given in the catalogs. According to these model the timescale, at which the SFR peaks, is around $3.5\lessapprox\tau\lessapprox 4.5$ [@speagleHighlyConsistentFramework2014]. If we intergrate the SFR(t), we can calculate the present day averaged SFR:

$$
\overline{\text{SFR}_\text{del}} = \frac{1}{t_{sf}}\int^{t_\text{end}}_{t_\text{start}}\text{SFR}_{del}(t)dt = \frac{A_{del}}{t_{sf}}\left[1-\left(1+\frac{t_{sf}}{\tau}\right)\exp\left(-\frac{t_{sf}}{\tau}\right)\right]
$$ {#eq-delayed-current-averaged}

Where $\overline{\text{SFR}}$ can be calculated by the stellar masses of the galaxy of interest:

$$
\overline{\text{SFR}_\text{del}} = \frac{\zeta M_*}{t_sf}
$$ {#eq-averaged-SFR-M*}

and $\zeta$ accounts for the mass-loss through stellar evolution, and is $\zeta \approx 1.3$ [@kroupaConstraintsStarFormation2020].

Using these two equations @haslbauerCosmologicalStarFormation2023 calculated the parameters $\tau,\ A_{del}$ of the model, for a constant $t_{sf}$ for all the galaxies, and thus they were able to approximate Cosmic SFRD of the LV. They found that the SFHs implied from the LV galaxies systematically underestimate the SFRDs at $z \lesssim 3$ and overestimates the SFRDs at $z\gtrsim 6$ as in comparison to those of MD14. In particular, the SFRD at the peak of the Lilly-Madau plot is $2.16 \pm 0.32$ lower in the case of the delayed-$\tau$. However, since they assume that all the galaxies begin at the same time and their sample size is not representative of the observed Universe, better calculations could potentially fix these discrepancies. However, these parametrization of the SFH could also be inadequate to describe galaxy evolution, especially for galaxies with $M_*<10^{10}\ M_\odot$, as discussed in @haslbauerCosmologicalStarFormation2023.

@@

To begin with the morphological types of the galaxies we first need to understand possible incompatibilities of the given types of each catalog.

::: {#fig-TTypes layout-ncol="2"}
![Before the clipping of the transient types](compare/quickplots/Hubble%20Morphology.png)

@@

Scatter plots and the residuals of the Hubble Numerical morphological types show a good correlation between the two datasets, if we ignore the transient galaxies.
:::
